\chapter{Risk Analysis and Ethics}
\label{chapter:ent-risk}

% Owner: Eesh

This chapter evaluates the risks facing the venture across technical, regulatory, commercial, and operational dimensions, stress-tests the financial projections under alternative scenarios, and addresses ethical and sustainability considerations.


\section{Scenario Analysis}
\label{sec:ent-scenarios}

\todo{Eesh: Draft this section.}


\subsection{Scenarios}

\todo{Define at least three scenarios:
\begin{itemize}
    \item \textbf{Base case:} as projected in \Cref{chapter:ent-financial}
    \item \textbf{Optimistic:} faster adoption, earlier regulatory clearance, stronger clinical evidence
    \item \textbf{Pessimistic:} regulatory delay, slower adoption, higher development costs, weaker clinical evidence
\end{itemize}
Present each scenario's impact on revenue, breakeven timing, and cash position. Clearly show which assumptions matter most commercially.}

\section{Risk Management}
\label{sec:ent-risk-mgmt}

\todo{Eesh: Draft this section.}

\subsection{Risk Register}

\todo{Present a risk register covering:
\begin{itemize}
    \item \textbf{Technical:} synchronisation limits between cameras, image quality under field conditions, false positive/negative rates, motion artefacts degrading measurement quality
    \item \textbf{Regulatory:} classification more stringent than expected, evidence requirements exceeding projections, \ab{samd}-specific obligations
    \item \textbf{Commercial:} buyer reluctance without strong evidence base, reimbursement uncertainty, workflow resistance from clinicians comfortable with existing tools
    \item \textbf{Operational:} supply chain fragility (Pi availability), manufacturing variability in optical alignment, key person dependency during early stages
    \item \textbf{Reputational/clinical:} overclaiming diagnostic capability, misuse by untrained operators, adverse outcome attributed to device result
\end{itemize}
For each risk: likelihood, impact, and specific mitigation.}

\subsection{Risk Mitigation Strategy}

\todo{Group mitigations thematically. Highlight which risks are addressed by design choices already made (e.g.\ quality gating, conservative claims language, local inference for data privacy) vs which require ongoing management.}


\section{Ethics, Sustainability, and Life-Cycle}
\label{sec:ent-ethics}

\todo{Eesh: Draft this section.}

\subsection{Ethical Considerations}

\todo{Address:
\begin{itemize}
    \item Concussion decisions affect real people's return-to-play and return-to-work outcomes --- consequences of false negatives and false positives
    \item Risk of over-reliance on device outputs by non-expert operators
    \item Inclusivity: has the system been validated across demographics, skin tones, eye colours, and physiological variability?
    \item Data governance and consent, particularly for minors in youth sport
    \item Algorithmic transparency: can the operator understand why the device gave a particular result?
\end{itemize}}

\subsection{Environmental Sustainability}

\todo{Address:
\begin{itemize}
    \item Battery and electronics disposal --- \ab{weee} compliance
    \item Material choice: PETG/PA12/polycarbonate lifecycle comparison
    \item Repairability and modular replacement (e.g.\ replaceable battery, facial interface, camera module) vs sealed disposable design
    \item Cloud vs local inference energy trade-offs --- local-first architecture reduces data centre energy use
    \item Packaging and shipping considerations
\end{itemize}}

\subsection{Product Life-Cycle}

\todo{Address:
\begin{itemize}
    \item Expected product lifetime and end-of-life strategy
    \item Upgrade path: can components be replaced incrementally or does the whole unit need replacing?
    \item Refurbishment and second-life potential for research or training use
    \item Responsible disposal and take-back programme
\end{itemize}}
