\chapter{Competitive Strategy and Defensibility}
\label{chapter:ent-competitive}

% Owner: Victor

This chapter sets out the headset's competitive strategy and defensibility, picking up where the market analysis of \Cref{chapter:ent-market} ends: given the \ab{mtbi} screening market characterised there, where does the headset fit, on what value-proposition basis, and behind what moats? It maps the competitive landscape, articulates the value proposition and product-market fit for each user segment, sets out the patent and non-patent defensibility position, and develops the willingness-to-pay evidence underlying a two-tier pricing architecture.


\section{Competitive Landscape}
\label{sec:ent-competitive-landscape}


\subsection{Direct Competitors}
\label{sec:ent-direct-competitors}

The commercial market for oculomotor screening divides into two tiers: portable but limited-accuracy devices, and accurate but inaccessible laboratory systems.

\subsubsection*{Portable, Limited-Accuracy Tier}

BrainEye (Neuro Optica Operations Pty Ltd, trading as BrainEye) is the closest commercial parallel. The current commercial product is a consumer smartphone screening app (iOS/Android, ${\sim}\pounds 40$/year subscription) that captures \ab{plr}, smooth pursuit, near point of convergence (\ab{npc}), and horizontal gaze nystagmus from the user's selfie camera. The most-cited validation reported 100\% sensitivity and 85\% specificity for concussion identification on a cohort of 348 adult professional male AFL footballers~\cite{clough2025braineye}, but that study was conducted not with the consumer app, but with a research hardware kit (fixed-distance cradle, chinrest, and overhead lightbar) required to impose the geometric and light control that the consumer form-factor cannot provide. The published accuracy figures therefore depend on a hardware kit the consumer app does not include. BrainEye and our headset address an overlapping biomarker set; our differentiation is not modality count but the controlled coaxial geometry of a head-mounted dichroic-routed optical path, which removes the user-positioning and lighting noise that the BrainEye consumer app inherits.

King-Devick (King-Devick Test, Inc.) is a rapid number-naming saccade test delivered as an iPad app or paper cards. Pricing is per-athlete tiered (${\sim}\$10$--$20$/athlete/year depending on cohort size, 3-athlete minimum); a typical school team of 25--30 athletes would pay ${\sim}\pounds 400$--$500$/year. No flat team-license tier is published. It indirectly assesses saccadic speed and accuracy but does not measure pupil light response or vergence. The test relies on clinician timing and is typically used as a screening adjunct rather than a primary tool.

Reflex (Brightlamp, Inc.) turns an iPhone into a pupillometer, measuring \ab{plr} in ${\sim}5$\,s using the camera flash; it is FDA Class~I, 510(k)-exempt. Pricing is a clinician subscription (${\sim}\$40$/month or \$400/year). It measures \ab{plr} only---no smooth pursuit, no vergence. A distinct but commonly confused product, EyeGuide Focus (EyeGuide Inc.), is a tabletop smooth-pursuit tracker stabilised by a chin rest; a 2025 elite rugby study found it did not discriminate concussed from non-concussed players, limiting its competitive threat. A separate commercial tier of concussion-management workflow platforms (Luca Health, Return2Play and similar) does not measure biomarkers but competes for the same institutional budget; these are discussed in \Cref{sec:ent-indirect-competitors}.

\subsubsection*{Accurate but High-Cost Tier}

EYE-SYNC (NeuroSync) is a VR-goggle-based system that measures smooth pursuit and saccadic eye movements with research-grade fidelity. The device costs approximately \pounds 25{,}000--30{,}000 and requires a trained operator to administer and interpret testing. (NeuroSync received an FDA warning letter in May 2025, introducing regulatory uncertainty around the product's current compliance status.) While accurate, the cost and workflow burden restrict use to elite sports medicine centers and university research programs.

Oculogica (EyeBOX) is a high-end binocular eye-tracking research system. It offers precise vergence and smooth pursuit measurement but operates at a similar price point and is designed for controlled laboratory environments, not field screening.

NeurOptics NPi-300 is a handheld automated pupillometer developed for critical care settings. It measures \ab{plr} with high repeatability (approximately $\pm 0.1$\,mm pupil diameter precision) and costs approximately \pounds 4{,}000 per unit (list price undisclosed; estimate based on NICE MIB235 advisory on the predecessor NPi-200). It provides no smooth pursuit or vergence capability and is not validated for concussion screening in athletic populations.

\begin{table}[htbp]
    \centering
    \renewcommand{\arraystretch}{1.5}
    \begin{tabular}{
        >{\raggedright\arraybackslash}p{2.4cm}
        *{3}{>{\centering\arraybackslash}p{1.15cm}}
        *{2}{>{\centering\arraybackslash}p{1.35cm}}
        >{\centering\arraybackslash}p{1.4cm}
        >{\raggedright\arraybackslash}p{1.9cm}
    }
        \toprule
         & \textcolor{pughheader}{\textbf{\ab{plr}}}
         & \makecell[c]{\textcolor{pughheader}{\textbf{Smooth}}\\\textcolor{pughheader}{\textbf{pursuit}}}
         & \textcolor{pughheader}{\textbf{Vergence}}
         & \textcolor{pughheader}{\textbf{Repeatability}}
         & \textcolor{pughheader}{\textbf{Portability}}
         & \makecell[c]{\textcolor{pughheader}{\textbf{Indicative}}\\\textcolor{pughheader}{\textbf{cost}}}
         & \makecell[l]{\textcolor{pughheader}{\textbf{Typical}}\\\textcolor{pughheader}{\textbf{setting}}} \\
        \midrule
        BrainEye (smartphone app)            & \hbfull & \hbfull & \hbfull & \hbnone & \hbfull & $\sim$\pounds 40/yr           & Consumer / parent \\
        King-Devick test                     & \hbnone & \hbhalf & \hbnone & \hbhalf & \hbhalf & $\sim$\pounds 400/yr (team)   & Clinical adjunct \\
        Reflex (Brightlamp)                  & \hbfull & \hbnone & \hbnone & \hbfull & \hbfull & $\sim$\pounds 400/yr          & Clinical \\
        NeurOptics NPi-300                   & \hbfull & \hbnone & \hbnone & \hbfull & \hbhalf & $\sim$\pounds 4k              & ICU / critical care \\
        EYE-SYNC (NeuroSync)                 & \hbnone & \hbfull & \hbfull & \hbfull & \hbnone & $\sim$\pounds 25--30k         & Elite sports medicine \\
        Oculogica EyeBOX                     & \hbhalf & \hbfull & \hbfull & \hbfull & \hbnone & $\sim$\pounds 25--30k         & Research laboratory \\
        \ab{scat6} / CRT6                    & \hbnone & \hbnone & \hbnone & \hbnone & \hbfull & Free                          & Sideline / clinical \\
        \bottomrule
    \end{tabular}
    \renewcommand{\arraystretch}{0.95}

    \smallskip
    {\small \hbfull~Capability present / high \hspace{1.5em}
            \hbhalf~Partial / qualified \hspace{1.5em}
            \hbnone~Absent / low}

    \caption{Competitor feature comparison across the major dedicated oculomotor and concussion-screening tools. Binary columns (\ab{plr}, smooth pursuit, vergence) mark whether the device measures the biomarker at all; repeatability and portability are scored qualitatively (high / partial / low). Indicative cost reflects the most-publicly-cited list or subscription price; see \Cref{tab:ent-price-benchmarking} for full sourcing. Our device is intentionally excluded from this table because its capability claims are pre-clinical-validation; its design-target position is shown separately in \Cref{fig:ent-positioning-matrix}. Concussion-management workflow platforms (Luca Health, Return2Play and similar) capture no biomarkers and are discussed separately in \Cref{sec:ent-indirect-competitors}.}
    \label{tab:ent-competitor-features}
\end{table}


\subsection{Indirect Competitors and the Status Quo}
\label{sec:ent-indirect-competitors}

Outside dedicated oculomotor devices, concussion screening relies on several lower-fidelity approaches.

The \ab{scat6} and its pediatric variant CRT6 are gold-standard clinical assessment protocols available free to practitioners. They comprise subjective symptom checklists, manual balance and coordination tests, and brief clinical questioning. Although clinically validated, published meta-analyses report pooled SCAT sensitivity in the 0.5--0.88 range depending on the component sub-test examined~\cite{dharnipragada2023scat}, with cognitive (SAC) and balance (m-BESS) sub-tests showing particularly degraded performance at 3--5 days post-injury~\cite{downey2018scat3}. The test is also vulnerable to symptom underreporting, particularly in young athletes who fear removal from play or judgment from peers.

Accelerometer-based mouthguards (e.g., Prevent Biometrics, VICIS) measure impact force during contact sport. They identify high-energy collisions but do not measure brain or neurological state. A player can sustain a severe impact force and show no concussion symptoms, or sustain a modest impact and suffer a clinically significant concussion. Impact magnitude is a risk factor but not a diagnostic marker.

Symptom-reporting apps (e.g., Heads Up, CDC concussion tracker apps) allow athletes to self-report symptoms over time. Self-report is vulnerable to social desirability bias---young athletes may underreport symptoms to remain eligible for play, particularly in communities where competitive success is especially valued.

Concussion-management workflow platforms occupy a distinct service tier in the same institutional buyer's budget. Luca Health (UK) is the most directly relevant: a B2B platform serving 10+ UK schools and grassroots clubs at the time of writing, combining smartphone-based incident logging, telehealth consultations with Luca-employed clinicians, return-to-play and return-to-school workflow management, and reporting back to clubs and schools. Pricing is per-athlete in the \pounds 15--20 per year range. Luca does not currently capture eye-movement or pupillometric biomarkers, so it is excluded from \Cref{tab:ent-competitor-features} (a feature comparison of measurement devices) and from \Cref{fig:ent-positioning-matrix} (whose axes are measurement-fidelity and access cost). Two features of its position nonetheless materially affect the medium-term competitive picture. First, Luca occupies a structurally privileged data position. Its CEO Nick Greenhalgh---a former Northampton Saints professional rugby player---characterised this directly in interview: \textit{``We're in a perfect position to collect data, basically simulating a clinical trial. We have the baseline, the post-concussion readings, and the true clinical assessment.''} Pre-injury baselines, post-injury readings, and ground-truth clinical assessments (delivered by Luca's own clinicians 24--48\,h later) are accumulated as a single longitudinal record per user---a dataset architecture whose defensibility advantage compounds over time. Second, Luca's product roadmap includes adding smartphone-based \ab{plr}, smooth pursuit, saccade and reaction-time testing to the consumer app; if delivered, this would place Luca in direct biomarker-overlap territory with the headset, although at the lower end of the measurement-fidelity axis (handheld form-factor, ambient light, no controlled NIR geometry). The headset's differentiation against Luca in that scenario remains the same as against BrainEye: controlled coaxial geometry, not modality count. Two adjacent UK workflow players---Return2Play (pitch-side physiotherapy for independent schools) and the Prime Practice (pitch-side first-aid)---occupy related niches without measurement capability.

\subsubsection*{The True Competitor: Status Quo}

In most grassroots youth sport settings---community football clubs, amateur rugby leagues, non-elite school sports---no objective screening tool is routinely used. Concussion detection relies on symptom recognition by untrained coaches, parental observation, or sideline judgment by team staff with minimal formal training. This status quo is the primary competitor for the headset.

Informal parental feedback (collected during preliminary design work) revealed significant demand but also cultural barriers. Mothers of young athletes expressed strong enthusiasm for objective screening, particularly when presented with evidence of cumulative concussion risks to neurodevelopment (``would definitely buy this'' was a common response). Fathers showed more divided attitudes: some equally supported investment in prevention, but a subset of parents with more ``tough love'' attitudes toward athletic injury were skeptical and viewed objective concussion screening as an unnecessary, protective overreach. This parental gender split reflects broader cultural variation in risk tolerance and sports safety norms within grassroots communities. The absence of any objective existing tool means that family attitudes toward concussion risk currently dictate screening practice. Objective and accessible measurement could address this, but only if adoption can overcome cultural resistance.


\subsection{Competitive Positioning Matrix}
\label{sec:ent-positioning-matrix}

\begin{figure}[htbp]
    \centering
    \begin{tikzpicture}[
        every node/.style={font=\scriptsize},
        comp/.style={circle, fill=black!85, inner sep=1.7pt},
        device/.style={dashed, very thick, draw=black},
    ]
        % Axes
        \draw[->, thick] (0,0) -- (10.2,0);
        \draw[->, thick] (0,0) -- (0,10.2);

        % Quadrant divider (subtle)
        \draw[dotted, gray!60] (5,0) -- (5,10);
        \draw[dotted, gray!60] (0,5) -- (10,5);

        % X-axis label and tick annotations
        \node[below, font=\small\bfseries] at (5, -1.4) {Cost / Inaccessibility $\rightarrow$};
        \node[below, font=\scriptsize] at (0.3, -0.15) {Free};
        \node[below, font=\scriptsize] at (3.0, -0.15) {$\sim$\pounds 400/yr};
        \node[below, font=\scriptsize] at (6.0, -0.15) {$\sim$\pounds 4k};
        \node[below, font=\scriptsize] at (9.0, -0.15) {$\sim$\pounds 25k+};

        % Y-axis label and tick annotations
        \node[rotate=90, font=\small\bfseries] at (-4.6, 5) {Measurement Fidelity $\rightarrow$};
        \node[left, font=\scriptsize, align=right, anchor=east] at (-0.15, 1.0) {Subjective /\\ symptom-report};
        \node[left, font=\scriptsize, align=right, anchor=east] at (-0.15, 3.0) {Limited objective\\ (1 modality,\\ low control)};
        \node[left, font=\scriptsize, align=right, anchor=east] at (-0.15, 5.5) {Single-modal,\\ controlled \emph{or}\\ multi-modal,\\ uncontrolled};
        \node[left, font=\scriptsize, align=right, anchor=east] at (-0.15, 8.5) {Multi-modal,\\ controlled\\ geometry};

        % Quadrant tags (italic, subtle)
        \node[font=\itshape\scriptsize, gray!75] at (1.5, 9.85) {empty quadrant};
        \node[font=\itshape\scriptsize, gray!75] at (7.5, 9.85) {clinical-grade};
        \node[font=\itshape\scriptsize, gray!75] at (1.5, 0.4) {consumer};
        \node[font=\itshape\scriptsize, gray!75] at (8.5, 0.4) {(no offering)};

        % Competitor points
        \node[comp, label={[font=\scriptsize]below right:SCAT6/CRT6}] at (0.3, 1.0) {};
        \node[comp, label={[font=\scriptsize]right:Symptom apps}] at (0.7, 1.6) {};
        \node[comp, label={[font=\scriptsize]left:King-Devick}] at (2.4, 2.4) {};
        \node[comp, label={[font=\scriptsize]right:Reflex}] at (3.7, 3.7) {};
        \node[comp, label={[font=\scriptsize]above right:BrainEye}] at (1.4, 5.4) {};
        \node[comp, label={[font=\scriptsize]above:NPi-300}] at (6.0, 5.5) {};
        \node[comp, label={[font=\scriptsize]below right:EYE-SYNC}] at (8.5, 7.8) {};
        \node[comp, label={[font=\scriptsize]left:Oculogica}] at (9.0, 9.0) {};

        % Our headset: dashed ellipse in the empty quadrant
        \draw[device] (2.5, 8.5) ellipse (1.6 and 0.9);
        \node[font=\bfseries\small] at (2.5, 8.5) {Our headset};
        \node[font=\scriptsize, align=center] at (2.5, 7.2) {(target position;\\ pre-clinical validation)};

    \end{tikzpicture}
    \caption{Competitive positioning chart for grassroots \ab{mtbi} oculomotor screening. Established competitors are plotted at their published cost and capability tiers. The dashed ellipse represents the headset's design target position: the high-fidelity claim derives from multi-modal capture (\ab{plr} + smooth pursuit + vergence) under controlled NIR illumination and head-mounted coaxial geometry (\Cref{tab:ent-traceability}; \Cref{chapter:headset}), rather than from clinical validation data, which is a Phase 1 deliverable (\Cref{chapter:ent-adoption-roadmap}). The low-cost, high-fidelity region is currently unoccupied by validated commercial devices.}
    \label{fig:ent-positioning-matrix}
\end{figure}

The competitive landscape reveals a striking gap (\Cref{fig:ent-positioning-matrix}). Devices that combine high measurement fidelity (simultaneous \ab{plr}, smooth pursuit, and vergence under controlled stimulus geometry) occupy the high-cost quadrant: EYE-SYNC, Oculogica, and NeurOptics span approximately \pounds 4{,}000--30{,}000. Devices that are cost-accessible (under \pounds 1{,}000) remain in the low-fidelity quadrant: \ab{scat6}, King-Devick, and symptom apps provide only indirect or subjective markers; BrainEye, as a smartphone app at ${\sim}\pounds 40$/year, sits closest to our consumer tier on price and captures an overlapping biomarker set, but delivers each measurement from a handheld form-factor whose positional and lighting variability undermine measurement fidelity. The low-cost, high-fidelity quadrant is empty. This gap exists because measurement fidelity, particularly for smooth pursuit and vergence, traditionally requires controlled optical geometry: fixed head position, coaxial stimulus and eye axis alignment, and stable working distance. Clinical systems achieve this through bulky headframes, gaze-tracking cameras mounted on optical tables, or VR goggle form factors that constrain portability. The beamsplitter architecture (\Cref{chapter:headset}) resolves this constraint by embedding coaxial geometry into a compact, wearable form factor. Precise stimulus and pupil alignment is achieved through optical design rather than mechanical positioning, allowing the device to deliver fidelity comparable to much larger systems at grassroots-accessible cost.


\subsection{Assumptions and Limitations}
\label{sec:ent-competitive-assumptions}

This competitive positioning rests on several assumptions that warrant explicit examination.

\textbf{Handheld Repeatability.} BrainEye is positioned as less repeatable than the headset, based on optical first principles (manual positioning introduces angle and distance variability that propagates into pupil measurement error). This claim is not empirically validated in a head-to-head comparison. Future work should benchmark the headset against BrainEye under controlled field conditions.

\textbf{Pricing Intelligence.} Estimates for NeuroSync/EYE-SYNC (\pounds 25{,}000--30{,}000), NeurOptics NPi-300 (${\sim}\pounds 4{,}000$), BrainEye (${\sim}\pounds 40$/year consumer subscription), and King-Devick (${\sim}\pounds 10$--$20$/athlete/year). Sources range from public product pages to NICE advisory reporting and distributor channels; actual UK list prices may vary and are not always publicly disclosed.

\textbf{Positioning Matrix Scope.} The matrix assesses measurement capability (what variables can each device measure, and with what precision). It does not reflect clinical validation status. Several devices in the high-fidelity quadrant (e.g., Oculogica, EYE-SYNC) are validated in research settings but may lack formal concussion-specific clinical trial data comparable to the \ab{scat6} or BrainEye's published athlete cohort study. Future clinical evidence for the headset will refine its position further.

\textbf{Market Definition.} This section focuses on devices explicitly designed for oculomotor or concussion screening. Broader medical devices (e.g., commercial eye-tracking for gaze-contingent virtual reality, research binocular eye trackers in neuroscience labs) exist at similar price points but operate in different markets with different regulatory and end-user expectations. The headset's competitive position assumes a primary market of grassroots sports medicine and youth athletic programs.


\section{Value Proposition and Product-Market Fit}
\label{sec:ent-value-prop}


\subsection{Target Users and Job-to-be-Done}
\label{sec:ent-target-users}

Three distinct user groups face concussion identification challenges in grassroots youth sport, each with a specific job-to-be-done that the headset addresses.

\subsubsection*{Group A: School PE Teachers}

School PE teachers are legally responsible for student welfare yet typically lack medical training. When a child sustains a head impact during sport, the teacher faces a critical decision: send the child home unsupervised, call parents immediately, contact emergency services, or allow continued participation. Current practice relies on verbal questioning and observation---sometimes supplemented by the \ab{scat6} questionnaire---but few teachers trust these subjective assessments to distinguish between minor and serious head injury. The core job-to-be-done articulated for this segment is: \textit{``give me confidence I'm making the right decision about whether this child needs medical attention, without requiring medical training.''} The liability-mitigation mechanism this JTBD implies---the reputational, financial and regulatory cost of a missed serious head injury at a school sporting event---was independently identified in interview by Greenhalgh as the primary commercial wedge with institutional decision-makers (\Cref{sec:ent-wtp-evidence}): \textit{``There's a willingness to pay from both schools and clubs, and it's driven by liability---they're terrified of a nurse missing a brain bleed and ending up in court.''} This corroboration is procurement-side rather than user-side, however: empirical validation of the JTBD directly with PE teachers remains a Phase 1 deliverable (\Cref{chapter:ent-adoption-roadmap}).

\subsubsection*{Group B: Grassroots Club Coaches}

Club coaches---particularly in rugby and football---operate with similar liability exposure but less institutional support. Weekend matches frequently occur without on-site medical staff or physiotherapists. Coaches report heightened awareness of concussion risk, driven by recent media coverage and athlete welfare campaigns. However, they face a dual pressure: protect player safety while avoiding unnecessary medical interventions that remove players from competition. The hypothesised job-to-be-done for this segment is: \textit{``confidently distinguish between a child who can safely return to play and one who needs medical evaluation, while protecting my club from liability.''} As with Group A, empirical validation through grassroots-coach interviews is a Phase 1 deliverable (\Cref{chapter:ent-adoption-roadmap}).

\subsubsection*{Group C: Parents}

Parents are the primary consumer segment, and the design has been calibrated against the parent persona below. After a child receives a head impact---whether at school, club, or recreational play---parents experience acute uncertainty: \textit{``Is this serious?''} The job-to-be-done for this segment is: \textit{``know right now whether my child needs medical assessment, so I can act with confidence.''} Sixteen semi-structured parent interviews (n = 9 mothers, n = 7 fathers) ground the persona that follows; \Cref{sec:ent-pmf-evidence} reports the segment-level willingness-to-pay signal these interviews surfaced.

\begin{tcolorbox}[
    colback=gray!4, colframe=gray!50,
    title={\textbf{Persona --- Sarah, 38, mother of two grassroots rugby players (aged 10 and 13)}},
    fonttitle=\small, fontupper=\small,
    boxrule=0.4pt, arc=2pt, left=5pt, right=5pt, top=4pt, bottom=4pt
]
\begin{description}[leftmargin=2.2cm, labelindent=0pt, style=sameline, itemsep=2pt, topsep=2pt, parsep=0pt]
    \item[\textbf{Motivations.}] Avoid missing a serious concussion; equally, avoid being the alarmist parent who pulls children out of fixtures unnecessarily.
    \item[\textbf{Behaviours.}] Watches school and club fixtures from the touchline; asks the coach after every visible knock; defaults to ``wait and see'' because A\&E is a two-hour round trip for a child who usually does shake it off.
    \item[\textbf{Experience goal.}] Feel like a responsible parent without being seen as overprotective.
    \item[\textbf{End goal.}] Within 5 minutes of a knock, know whether the child needs medical attention or can safely return to play.
    \item[\textbf{Pains.}] Low trust in self-report (\textit{``he says he's fine but\dots''}); no objective tool available at the touchline; persistent concern about cumulative damage from sub-clinical hits.
    \item[\textbf{WTP framing.}] \textit{``If I could use something like this after a bad knock, I'd know immediately whether to take them to hospital. I'd definitely buy this---I worry constantly about repeat concussions affecting their development.''} Frames the purchase alongside car seats, helmets, and first-aid training rather than alongside clinical diagnostics.
\end{description}
\end{tcolorbox}

The parent segment is not monolithic. One father of a 12-year-old rugby player expressed skepticism (\textit{``kids are tougher than people think''}), representing a ``tough love'' sub-segment unconvinced that the risk warrants adoption. The marketing implication is that adoption messaging must emphasise evidence-based developmental risk rather than general safety rhetoric, in order to convert this sub-segment rather than alienate it.


\subsection{Design Requirements Traceability}
\label{sec:ent-traceability}

The headset's competitive advantage emerges from systematic alignment between user needs, measurement requirements, and hardware specifications. \Cref{tab:ent-traceability} traces this chain for five core requirements.

\begin{table}[htbp]
    \centering
    \caption{Design Requirements Traceability --- left-to-right chain linking each user need to the corresponding measurement requirement, hardware specification, and resulting competitive advantage versus existing solutions (\ab{scat6}, EYE-SYNC, BrainEye, clinical eye-trackers).}
    \label{tab:ent-traceability}
    \begin{tabular}{p{2.7cm}p{3.0cm}p{3.4cm}p{3.4cm}}
        \toprule
        \textbf{User Need} & \textbf{Measurement Requirement} & \textbf{Hardware Spec} & \textbf{Competitive Advantage} \\
        \midrule
        Non-medical user requires reliable objective output
        & No subjective interpretation; controlled measurement geometry
        & Automated pupillometry + smooth pursuit + vergence tracking with automated thresholding
        & Objective + geometry-controlled output; unlike \ab{scat6} (subjective) or BrainEye (handheld form-factor) \\
        \addlinespace
        Sideline use: rapid deployment at match without infrastructure
        & Portable, self-contained assessment within 90\,s without external equipment
        & Head-mounted with integrated display + NIR camera; battery powered; 3D-printed enclosure; low-cost component selection
        & Field-deployable; unlike EYE-SYNC (tethered to stationary computer) or Saccadometer (requires clinical lab setting) \\
        \addlinespace
        Safe for children aged 8--18; no risk of phototoxicity
        & Comfortable fit, scalable; no eye damage
        & Plano-convex BK7 singlet, $f = 100$\,mm; 850\,nm NIR; IPD range 48--65\,mm; IEC 62471 safe exposure
        & Child-safe optical power; unlike adult-only devices (EyeLink, Tobii); adjustable IPD supports varied anatomy \\
        \addlinespace
        Affordable for school/club budget (competing with \ab{scat6}, which is free but subjective)
        & Cost-constrained design within consumer-tier price band
        & Off-the-shelf optics: longpass dichroic plate beamsplitter, plano-convex BK7 singlets, NIR LED \& Digi-Key; commodity compute and battery
        & Low cost vs.\ clinical devices (\pounds 15k--25k); vs.\ existing screening tools (free but subjective) \\
        \addlinespace
        Defensible decision-making (coach/teacher liability)
        & Rapid, documented, objective output (\ab{plr}, SP asymmetry, vergence) with timestamp
        & Automated analysis; on-device data logging; timestamped results with pass/refer thresholds
        & Audit trail + objective evidence; reduces coach/teacher liability vs.\ verbal-only assessment \\
        \bottomrule
    \end{tabular}
\end{table}

This traceability demonstrates that the headset is not a technology solution seeking a problem. Rather, each engineering choice (beamsplitter optics, 850\,nm NIR, automated pupillometry, 3D-printed enclosure) directly addresses a user need identified through the job-to-be-done framework. The competitive advantage accrues precisely because the headset occupies a market quadrant---\textit{high accuracy, low cost, field-deployable, operator-agnostic}---left empty by existing tools.


\subsection{Evidence of Product-Market Fit}
\label{sec:ent-pmf-evidence}

Current practice in concussion screening at grassroots level relies on three approaches: (1) no formal assessment, relying entirely on player self-report; (2) \ab{scat6} questionnaire, which is free and evidence-based but requires subjective interpretation and medical training; or (3) referral to hospital A\&E, which is costly, time-consuming, and unnecessary for minor impacts. None of these options provides what users actually need: \textit{rapid, objective, operator-agnostic screening at the point of injury}.

Parent interviews reveal strong product-market fit signals within this segment. All mothers interviewed ($n = 9$) who learned about the oculomotor screening concept expressed high purchase intent, particularly when presented with evidence linking repeated concussions to child cognitive and developmental risk. One mother stated unprompted: \textit{``I would definitely buy this.''} Fathers showed more heterogeneity ($n = 7$ interviewed): one expressed equal enthusiasm, while another remained skeptical of the underlying risk premise, highlighting that parental adoption will require evidence-based messaging about developmental consequences rather than general safety rhetoric.

Critically, parents demonstrated willingness to pay. Informal discussions suggested acceptance of pricing in the \pounds 250--\pounds 300 range, supporting the consumer pricing structure outlined in \Cref{sec:ent-pricing-strategy-victor}. This sits well below clinical alternatives (\pounds 4k--\pounds 30k), consistent with Christensen's non-consumption / new-market entry framing~\cite{christensen2003innovatorssolution}: the device competes against the status quo of grassroots screening rather than seeking to displace clinical incumbents.

Teacher and coach feedback remains preliminary; both groups expressed theoretical interest in objective assessment but have not yet been formally surveyed on adoption likelihood, price sensitivity, or integration barriers. Formal interviews with school PE professionals (focusing on liability-mitigation framing) and grassroots-club coaches (focusing on pitch-side workflow integration) are central activities of Phase 1 stakeholder research (\Cref{chapter:ent-adoption-roadmap}); the product-market-fit claims for these segments are honestly preliminary until that work is complete.


\subsection{Assumptions and Limitations}
\label{sec:ent-pmf-assumptions}

Three key assumptions underpin the product-market fit narrative and should be validated during user research.

First, the analysis assumes that grassroots schools and clubs will adopt an objective screening tool despite its lack of full clinical validation (gold-standard validation remains in progress). This assumption is reasonable given the weakness of current alternatives (\ab{scat6} subjectivity, no sideline access to medical imaging), but adoption rates remain unproven.

Second, parental enthusiasm was elicited during informal conversations with a small, UK-based sample; selection bias may inflate stated purchase intent. Formal quantitative validation (e.g., conjoint analysis or pre-orders) is required before claiming strong market demand.

Third, the analysis is geographically UK-centric and reflects English school PE and grassroots club structures. Applicability to other education systems, sports cultures, or countries requires separate validation. The regulatory environment for medical devices also varies materially across jurisdictions, which may affect market access timelines and adoption costs.

Fourth, the relative weighting of consumer (Group C) versus institutional (Groups A and B) demand is contested. The Group C willingness-to-pay signal---parent purchase intent in the \pounds 250--300 range---is grounded in 16 semi-structured parent interviews. A contrary read emerged in the Greenhalgh interview (\Cref{sec:ent-wtp-evidence}): \textit{``I'm not convinced the consumer market is large enough. Institutions are already doing so much on concussion---policies, nursing teams, awareness training---and parents aren't going to trust this kind of device without formal diagnostic status.''} Combined with the rarity of paediatric concussion events at the individual-family level, this suggests the residual consumer market may be both smaller and harder to reach than parent interviews alone would suggest. The challenge does not invalidate the consumer thesis, but it shifts the burden of proof: resolving the relative consumer-vs-institutional weighting requires the formal quantitative methods identified in assumption two (conjoint analysis, pre-orders) combined with direct procurement-side interviews with school and club buyers, both of which are Phase 1 deliverables (\Cref{chapter:ent-adoption-roadmap}).

These limitations do not undermine the value proposition, but they bound the scope of claims in this section and identify critical risks for mitigation during the validation phase.


\section{IP Strategy and Defensibility}
\label{sec:ent-ip}

Patent protection is one of three complementary defensibility levers the headset relies on; the other two---regulatory certification and biometric-data governance for under-18 users---are developed in \Cref{chapter:ent-regulatory} and \Cref{chapter:ent-risk} respectively. This section restricts itself to the patentability of the optical architecture, and treats certification and data-governance moats as cross-references rather than independent claims.


\subsection{Patentability of the Optical Configuration}
\label{sec:ent-patentability}

Dichroic spectral routing for eye imaging is densely populated prior art rather than a thin claim space, and the patent moat must be sized accordingly. \citet{taboada1994} anticipates the polarity-flipped version of our spectral routing (NIR reflected to the eye, visible transmitted to the camera) as far back as 1992; \citet{neuropticspatent2017} covers dichroic-routed \ab{plr} measurement for handheld pupillometry; HMD-side dichroic eye tracking is owned by Apple, Meta, 3M and Magic Leap across multiple grants between 2014 and 2024~\cite{applewaveguide2023, mmreyetrack2018, oculusdichroic2019}. Multi-modal oculomotor concussion screening in a portable headset is anticipated by \citet{neurokineticshmd2017}, \citet{krueger2014ocular} and the SyncThink/Neurosync portfolio (16+ grants, FDA-cleared as a concussion-aid HMD)~\cite{syncthinkpatent2015}. Multi-modal low-cost screening combining \ab{plr}, smooth pursuit and convergence is publicly disclosed in BrainEye~\cite{clough2025braineye} and FOVE-VR pupillometer studies~\cite{aleci2025fove}---non-patent prior art under \S102 / Art.~54 EPC. The defensible aperture is therefore not the individual elements, but a narrow conjunction.

The narrowest defensible filing target identified by this scan is a dependent claim reciting the specific opto-mechanical conjunction of [longpass dichroic at 750\,nm cut, plano-convex BK7 relay singlet $f \approx 100$\,mm, 850\,nm NIR ring illumination, septum-divided binocular display, and a multi-modal signal-fusion method operating on simultaneously-acquired \ab{plr}, smooth-pursuit, and vergence streams], conditional on patent counsel identifying a non-obvious technical synergy that the conjunction solves---for example, an alignment-tolerance specification it achieves at low cost, a calibration scheme that survives sideline mechanical shock, or a specific multi-modal signal-fusion algorithm. Without such a synergy, the conjunction faces moderate-to-high \S103 / Art.~56 obviousness risk. Cost-of-goods positioning is excluded from the patentability argument: it is not patentable subject matter under any office's examination guidelines and operates as commercial differentiation only.

A US prosecution strategy must additionally navigate \S101 / \emph{Mayo}--\emph{Alice} exposure for screening-method claims: \emph{PerkinElmer, Inc.\ v.\ Intema Ltd.}~(Fed.\ Cir.\ 2012) held a screening-method claim ineligible under the abstract-idea framework, and screening framing does not avoid \S102 or \S103 rejections, which are technology-feature-based rather than application-name-based~\cite{perkinelmerintema2012}. The screening-aid positioning that strengthens regulatory classification (\Cref{chapter:ent-regulatory}) does not strengthen patent eligibility and, in the US, may worsen it. Filing within 12 months of first disclosure secures Paris Convention Art.~4 priority while prototype validation continues~\cite{pariscv1883}. \textbf{Of the three complementary defensibility levers introduced at the start of \Cref{sec:ent-ip}, patents are the weakest and slowest pillar: the bulk of the commercial moat is carried by the regulatory-certification timeline barrier (\Cref{chapter:ent-regulatory}) and the biometric-data-governance regime governing longitudinal calibration data accumulated during pilot deployment (\Cref{chapter:ent-risk}), each developed in subsequent chapters.} A full prior-art search methodology, including search engines, query strings, classification codes, date range, and coverage limits, is documented in \Cref{appendix:prior-art-search}.




\section{Pricing Strategy and Willingness-to-Pay}
\label{sec:ent-pricing-strategy-victor}


\subsection{Competitor Price Benchmarking}
\label{sec:ent-price-benchmarking}

The oculomotor screening market spans three distinct price tiers, each reflecting different buyer segments and deployment contexts. High-end clinical systems (EYE-SYNC, ${\sim}\pounds 25$--$30$k; NeurOptics NPi-300, ${\sim}\pounds 4$k) target hospital and intensive care settings where capital budgets support substantial equipment investment and regulatory compliance justifies premium pricing. The consumer tier is dominated by smartphone-based SaaS (BrainEye ${\sim}\pounds 40$/year consumer subscription; Reflex ${\sim}\pounds 400$/year clinician subscription) and per-athlete licensing (King-Devick ${\sim}\pounds 10$--$20$/athlete/year). \ab{scat6} is paper-based and free. Our device targets a critical market gap---multi-modal \ab{plr} + smooth pursuit + vergence screening at consumer-accessible prices---where no established competitor currently operates. The price bands implied by the willingness-to-pay analysis below frame the unit economics developed in Rahul's financial model (\Cref{chapter:ent-financial}).

\begin{table}[htbp]
    \centering
    \caption{Competitor price benchmarking across the \ab{mtbi} screening market.}
    \label{tab:ent-price-benchmarking}
    \begin{tabular}{p{3.0cm}p{2.6cm}p{2.4cm}p{2.4cm}p{2.4cm}}
        \toprule
        \textbf{Device} & \textbf{Price} & \textbf{Target buyer} & \textbf{Model} & \textbf{Source} \\
        \midrule
        EYE-SYNC (NeuroSync) & ${\sim}\pounds 25$--$30$k & Hospital/clinic & Capital & Public reporting \\
        \addlinespace
        NeurOptics NPi-300 & ${\sim}\pounds 4$k & ICU & Capital & NICE MIB235 (NPi-200 predecessor) \\
        \addlinespace
        BrainEye & ${\sim}\pounds 40$/yr & Parents/prosumer & SaaS subscription & App Store \\
        \addlinespace
        King-Devick & ${\sim}\pounds 10$--$20$/athlete/yr & Coach/clinician & Per-athlete subscription & Public \\
        \addlinespace
        \ab{scat6} & Free & Anyone & Paper & N/A \\
        \addlinespace
        Our device (consumer) & \pounds 250--300 ceiling (parent WTP) & Parents & Outright purchase & \Cref{sec:ent-wtp-evidence}; pricing in \Cref{chapter:ent-financial} \\
        \addlinespace
        Our device (institutional) & Mid-hundreds hardware + recurring service fee & Schools, teams & Hardware + subscription & \Cref{sec:ent-wtp-evidence}; pricing in \Cref{chapter:ent-financial} \\
        \bottomrule
    \end{tabular}
\end{table}



\subsection{Willingness-to-Pay Evidence}
\label{sec:ent-wtp-evidence}

\textbf{Consumer segment (parents).} Interviews with mothers revealed strong appetite to purchase at \pounds 200--300 for objective concussion screening at touchline or home; parents explicitly framed the purchase alongside other child safety investments (car seats, cycling helmets, first aid training) rather than comparing to clinical diagnostics. This signals acceptance of the device as a preventive safety tool rather than a medical instrument. Fathers exhibited more heterogeneity: some equally enthusiastic, others skeptical of perceived overprotection, yet none rejected the concept outright based on pricing alone. Consumers will not pay above \pounds 250--300. No consumer subscription is planned---the core screening function is always available without restriction.

\textbf{Institutional segment (schools, teams).} Schools and teams represent higher willingness-to-pay than consumers (mid-hundreds for hardware, potentially approaching \pounds 1{,}000) because the device mitigates institutional liability and duty-of-care obligations. Institutional buyers additionally exhibit willingness to pay for recurring services that consumers do not: (1) on-site baseline administration---a trained operator visits the school to establish per-student baseline oculomotor measurements, improving screening accuracy relative to population-norm comparison; (2) longitudinal data storage---individual student records tracked over time, supporting return-to-play decisions and parental reporting; (3) aggregate statistics---school-level concussion risk data, supporting safeguarding reports and insurance requirements. This dual willingness---one-off hardware plus recurring service fee---supports a hybrid pricing architecture; the specific subscription pricing, unit economics, and recurring-revenue dynamics are developed in Rahul's financial model (\Cref{chapter:ent-financial}). A targeted procurement-side interview with Greenhalgh independently corroborated three elements of the institutional case. First, the liability-mitigation mechanism---the reputational, financial and regulatory cost of a missed serious head injury at a school sporting event---was identified as the primary commercial wedge with institutional decision-makers, corroborating the liability-driven JTBD framing of \Cref{sec:ent-target-users}. Second, Luca's own commercial model---a per-athlete annual subscription at \pounds 15--20 per year---provides a comparable institutional benchmark: at typical UK contact-sport-team headcounts (25--30 athletes), this implies \pounds 375--600 per team per year, falling within the per-device subscription band developed in \Cref{chapter:ent-financial}. Greenhalgh added that schools negotiate this cohort actively: \textit{``Schools try to be savvy. They ask: do we really need to sign up the whole school? Some narrow it to just the rugby team, to save money.''} Third, two procurement-side frictions bound the case: acute budgetary sensitivity in the current VAT-pressure environment, and internal-political resistance from school medical staff who perceive external concussion services as undermining their role. As Greenhalgh put it: \textit{``The objection often comes from the school's medical centre---they feel put out, put on the backburner.''} The latter friction is materially mitigated, in our case, by the headset's point-of-impact, coach- or teacher-operated workflow, which augments rather than replaces in-house medical capacity. These estimates remain directional; full procurement-side validation through direct school and club interviews is a Phase 1 deliverable (\Cref{chapter:ent-adoption-roadmap}).

\textbf{Ethical constraint.} The device always performs core screening---determining whether someone is at increased risk of concussion---without restriction. The concussion risk assessment is never paywalled. Subscription services cover data, tracking, and baseline administration. This distinction is critical: paywalling a safety-critical screening function would be ethically indefensible and commercially damaging.


\subsection{Pricing Model Evaluation}
\label{sec:ent-pricing-model}

The willingness-to-pay evidence above implies a differentiated pricing architecture by segment, not a single price point.

\textbf{Consumer segment.} The \pounds 250--300 ceiling and parents' explicit framing of the purchase alongside discretionary safety investments (car seats, helmets, first-aid training) constrains the consumer offering to a one-off outright purchase. A consumer subscription would compound poorly against a one-off-purchase mental model and risk losing adoption to subscription fatigue, particularly for a device whose value is concentrated at the rare moment of injury rather than in continuous use.

\textbf{Institutional segment.} The combination of (i) higher absolute willingness-to-pay for hardware (mid-hundreds approaching \pounds 1{,}000) and (ii) demonstrated willingness to pay separately for ongoing services (baseline administration, longitudinal records, aggregate safeguarding statistics) supports a hybrid model: hardware sale plus a recurring subscription tied to the service layer rather than to core screening functionality. This dual revenue structure is what the demand-side evidence sustains; the specific subscription pricing, unit economics, break-even volumes, and scenario analysis are developed in Rahul's financial model (\Cref{chapter:ent-financial}).


\section{Commercial Limitations}
\label{sec:ent-commercial-limitations}

Three limitations bound the demand-side claims of this chapter. First, the primary research sample---informal conversations with parents of young athletes and preliminary engagement with grassroots practitioners---provides directional insights into screening demand but lacks statistical power to quantify adoption barriers or willingness-to-pay distributions across diverse cohorts (different sports, age groups, socioeconomic contexts). Formal validation methods (conjoint analysis, pre-orders) are required before treating the WTP bands above as anything more than directional. Second, the regulatory pathway remains preliminary; while Class IIa medical device classification under MDR (active, non-invasive, screening-aid purpose) is the probable route (developed in \Cref{chapter:ent-regulatory}), formal pre-submission engagement with \ab{mhra} or a notified body has not commenced. Third, adoption velocity in the institutional channel is a structural constraint. UK schools are recognised in the sector as slow-moving institutional buyers, with sales cycles routinely spanning multiple academic-term boundaries. Beyond procurement-cycle latency, both schools and grassroots clubs commonly perceive existing concussion-policy frameworks---HEADCASE training, the UK Government's Grassroots Concussion Guidelines, in-house nursing teams---as already adequate, requiring a market-education phase to precede any meaningful sales conversation. Greenhalgh framed this directly: \textit{``The clubs already think they're doing enough. We have to convince them otherwise before we can sell anything.''} This pattern is consistent with the non-consumption / new-market entry framing already adopted in \Cref{sec:ent-pmf-evidence}, and motivates a top-down channel strategy: partnership with governing sporting bodies and educational associations (e.g.\ the RFU, the Medical Officers of Schools Association, and England Rugby Schools) rather than direct school-by-school sales, developed further in the adoption roadmap (\Cref{chapter:ent-adoption-roadmap}).

The four threads of this chapter---competitive landscape, value proposition, IP defensibility, and pricing architecture---converge on a single strategic position: a screening-aid device that fills an empty quadrant in the competitive map, addresses a documented job-to-be-done for each user group, defends an optical-architecture moat, and supports a pricing architecture grounded in segment-specific willingness-to-pay. The remaining strategic question---\textit{who adopts first, and in what sequence}---is the subject of the adoption roadmap and channel sequencing developed in \Cref{chapter:ent-adoption-roadmap}.
