\chapter{Market Analysis}
\label{chapter:ent-market}

% Owner: Suleiman (market segmentation, stakeholders) + Victor (value proposition, PMF)


This chapter analyses the market opportunity for a headset-based oculomotor screening system for mild traumatic brain injuries (mTBI). It first looks into the broader mTBI screening market including sport, medical and clinical applications, then focuses on the market of children and adolescents in grassroots sports that the product is aimed towards. Additionally, this chapter answers why this market was chosen and quantifies the opportunity through the Total Addressable Market (TAM), Serviceable Addressable Market (SAM) estimates and segments the market by deployment setting: youth sport clubs, schools and home use.  


\section{Market Segmentation}
\label{sec:ent-market-analysis}

\subsection{The \ab{mtbi} Screening Market}

The global \ab{mtbi} assessment market was valued at \$1.17 billion in 2025, with a valuation of \$3.91 billion by 2035 at a CAGR of 12.7\%~\cite{businessresearchinsights2025tbi}. This market spans multiple domains such as sports (professional, amateur and youth), military (through explosions), clinical settings (A\&E) and occupational health.

Within sport, football is the world's largest participation sport with 265 million total players globally~\cite{fifa2007bigcount}, while the same FIFA source reports 21.5 million registered youth footballers under 18. Rugby union accounts for 8.46 million players globally~\cite{worldrugby2023participation}, with 57\% reported as pre-teen, and has one of the highest concussion rates per exposure where in elite rugby, it is 4.73--9.05 per 1000 exposures~\cite{gardner2014rugbyconcussion,rafferty2019rugbyconcussion} among the mainstream contact sports, with others such as basketball, hockey and martial arts contributing to this demographic.

\subsection{Market Focus}

The specific market focused within this project is non-diagnostic \ab{mtbi} screening for children and adolescents in grassroots sport, including school rugby, school football, youth clubs and amateur leagues. Selecting this as the primary beachhead is justified through infrastructure, ethical and behavioural economic drivers which can create a rich environment for adoption.

\paragraph{Infrastructure.} Within grassroots settings, it is reported there are 0.2 to 0.5 per 1000 athlete exposures but this is a severe undercount of the reality due to the lack of trained medical staff in these situations to do methods such as SCAT6~\cite{echemendia2023scat6}. In grassroots settings, reliance upon CRT6 by untrained volunteers can create a large safety gap with the data showing that if a concussion is only managed by a coach using CRT6 then the injury is missed in 42\% of cases compared to when an on-site trainer does CRT6~\cite{echemendia2023crt6}.

\paragraph{Ethics.} There is a larger societal duty of care towards the adolescent brain which is still developing, exhibiting higher vulnerability to Second Impact Syndrome (SIS) and prolonged recovery times in comparison with adults~\cite{patricios2023amsterdam}. This can be a primary driver for adopting screening technologies.

\paragraph{Behavioural economics.} Adoption in this segment can also be catalysed through self-other discrepancy in risk perception~\cite{stone2013selfother,dore2014parental}. Individuals can exhibit cognitive biases which lead to the acceptance of personal physical risks but studies show that they demonstrate much higher levels of risk-aversion when making safety decisions for others. The behaviour driver is rooted in Social Values Theory suggesting that parents and teachers will follow social norms to make safe decisions for children in their case. These combinations of safety biases and the institutional duty of care act as powerful adoption catalysts which will help reduce the market friction for the tool.
\subsection{Settings Segmentation}

\paragraph{Youth sports clubs.} Youth sports clubs are a key area of interest for deploying the tool. Across Europe it is estimated that there are 700,000 sports clubs~\cite{nagel2020europeansportsclubs}, creating a broad club infrastructure through which an EU/EEA commercial launch can reach children in organised grassroots sport. Football offers the widest route through UEFA-affiliated school-to-club pathways, while the EU Six Nations rugby subset in \Cref{tab:ent-six-nations-rugby} provides a complementary contact-sport route. UK and England figures are retained only as validation comparators rather than the paid commercial beachhead.

\paragraph{Schools.} Schools are the most structured institutional route for an EU/EEA launch because PE teachers and school leaders already coordinate sport participation, safeguarding procedures and parent communication. UEFA's Football in Schools programme has engaged 41,500 schools and introduced 7.2 million children to football across its member associations~\cite{uefa2025schools}. Eurostat's older 93.3 million EU education figure includes tertiary and post-secondary education~\cite{eurostat2022students}; a more age-appropriate school-age frame is primary to upper-secondary education, where EU enrolment is approximately 60 million pupils across ISCED 1--3~\cite{eurostat2023teachers,eurostat2025secondary}. Schools would purchase the screening tool for usage in PE and sport fixtures.

\paragraph{Homes.} This is for children who come home from schools with concussions as well as parents who are concerned about their children's wellbeing playing grassroots sports. Parents would purchase the screening tool for use at home and for their child to take to school in high-risk sports-related activities. Traditionally, parents would use high variance sources such as Google as their first steps in this situation~\cite{kubb2020parentshealthinfo} whereas this screening tool can provide a first step of low variance objectivity to assist in this process.

\section{Market Sizing: TAM and SAM}

The Total Addressable Market (TAM) is treated here as a bounded opportunity rather than a single exact figure. For the upper bound, UNESCO's global school enrolment estimates that there are 1.4 billion children, adolescents and youth in primary and secondary schools~\cite{unesco2026gem}, providing a theoretical ceiling. If all of these households were addressable with the headset costing \pounds{}300, this maximum global TAM would be \pounds{}420 billion. However, because not all of these children will have the same exposure to sport-related concussions, a more conservative lower bound with stronger evidence can be made: FIFA reports 21.5 million registered youth footballers under 18~\cite{fifa2007bigcount}, and World Rugby reports 8.46 million rugby participants globally, 57\% of whom are pre-teen~\cite{worldrugby2023participation}, giving approximately 4.8 million youth/pre-teen rugby participants. These two sports alone imply 26.3 million high-relevance youth users equivalent to a conservative TAM floor of \pounds{}7.9 billion, before including unregistered football, school sport, hockey, basketball, martial arts or other head-impact settings. This lower bound is used without including numerous other sports to ensure precision, trading the limitations of the number of sports for the high accuracy of these citations.

The Serviceable Addressable Market (SAM) narrows down to school children in Europe that can be served after UK/Oxford validation, CE-marked market access and institutional case studies. This is smaller than the global TAM because initial routes to adoption rely on trusted schools, clubs and parent-facing institutions rather than direct global consumer sales, as evidenced in \Cref{chapter:ent-gtm}. The first commercial SAM is defined around the EU/EEA school-football channel reachable after CE-marked launch, using UEFA Football in Schools as the nearest scaled route to schools, teachers and school-to-club pathways. The programme has introduced 7.2 million children to football and engaged 41,500 schools across Europe~\cite{uefa2025schools}. At \pounds{}300 per parent device, this implies a \pounds{}2.16 billion parent opportunity; the institutional opportunity from 41,500 schools at \pounds{}2,700 per first-year institutional package adds \pounds{}112.05 million. The resulting European school-football SAM is approximately \pounds{}2.27 billion. This remains conservative because it excludes wider grassroots clubs, unregistered sport, rugby, hockey, basketball, martial arts, cycling and home-use demand outside organised football.

Further geographic segmentation can be done here to show why the EU/EEA launch can extend beyond school football into rugby. \Cref{tab:ent-six-nations-rugby} is an illustrative European rugby comparator using one source base of the World Rugby participation-map~\cite{worldrugby2020participationmap}. Because England, Wales and Scotland are not EU/EEA commercial-launch countries, the table includes an EU subset subtotal for France, Ireland and Italy.

\begin{table}[htbp]
    \centering
    \footnotesize
    \caption{Illustrative Six Nations rugby participation view.}
    \label{tab:ent-six-nations-rugby}
    \begin{tabular}{lrrrr}
        \toprule
        \textbf{Country} & \textbf{Registered players} & \textbf{Est. youth (<18)} & \textbf{Youth \%} & \textbf{Clubs} \\
        \midrule
        England     & 1,990,988 & 1,094,043 & 55\% & 1,809 \\
        France      &   542,242 &   244,009 & 45\% & 1,874 \\
        Ireland     &   101,922 &    50,961 & 50\% &   213 \\
        Italy       &    87,211 &    39,245 & 45\% &   490 \\
        Wales       &    83,120 &    41,560 & 50\% &   357 \\
        Scotland    &    49,265 &    24,633 & 50\% &   251 \\
        \midrule
        EU subset (France, Ireland, Italy) &   731,375 &   334,215 & 46\% & 2,577 \\
        Six Nations total & 2,854,748 & 1,494,451 & 52\% & 4,994 \\
        \bottomrule
    \end{tabular}
\end{table}

The EU Six Nations subset exposes approximately 334,000 youth players across 2,577 clubs, supporting EU/EEA rugby pilots without relying on England as the commercial launch evidence. The full Six Nations table is retained as a European comparator because it shows the broader contact-sport opportunity around the initial school-football SAM.

\section{Conclusion}
This chapter has located a segment where there is a need for mTBI screening and bounded the commercial opportunity around it. The TAM is treated as a \pounds{}7.9 billion conservative global floor from youth football and rugby participation alone. The first commercial SAM is approximately \pounds{}2.27 billion for the European school-football channel, formed from \pounds{}2.16 billion of parent opportunity and \pounds{}112.05 million of institutional school opportunity. This excludes wider grassroots clubs, rugby, unregistered sport, hockey, basketball, martial arts and home-use demand outside organised football. Youth grassroots sports was shown to be a priority segment through the infrastructure gap in CRT6 administration~\cite{echemendia2023crt6}, larger vulnerability to second impact syndrome~\cite{patricios2023amsterdam}, and the self-other risk discrepancy that orients parental decision-makers towards safety-first behaviour~\cite{stone2013selfother,dore2014parental}. The three deployment settings including schools, homes and youth sports clubs exhibit buyer dynamics that the go-to-market strategy in \Cref{chapter:ent-gtm} will build on.
