\chapter{Go-to-Market Strategy}
\label{chapter:ent-gtm}

% Owner: Suleiman

This chapter develops the go-to-market execution strategy for commercialising the headset-based mTBI screening system, building on the adoption sequence established in \Cref{chapter:ent-adoption-roadmap}. Phase 1 remains a UK research-validation phase; paid institutional rollout begins in the EU/EEA commercial beachhead after CE-marked market access. The first section uses that sequence to define how each phase is executed through channels, pilot design, timelines and phase-gate metrics. The final section then converts this plan into a Serviceable Obtainable Market (SOM), using the number of institutions reached through the GTM plan.


\section{GTM Execution Strategy}
\label{sec:ent-gtm-execution}

\Cref{chapter:ent-adoption-roadmap} outlined the order of the GTM strategy and the reasons for that order. In this section, the exact methods of how these phases are executed include: the channels used, the pilot design, the timeline, the offer structure and the evidence required before progressing. User interviews and user journeys will be used again to identify points of adoption friction so that decisions are made with the intention of reducing them.

Each GTM phase in this section is treated as an execution experiment rather than a fixed launch strategy. Since insufficient deployment data has yet to be collected, KPIs are used as phase-gate metrics. These are used to test whether our choices in channels, messages and deployment models are working well enough to justify scaling within the phase or progressing onto the next phase. If the KPI metrics are not met, then iterations must occur within the phase instead of progression. This dynamic, iterative approach provides exceptional flexibility to ensure that we do not get stuck in performing sub-optimal tactics and instead discover flaws in smaller units of a strategy before expanding it.

The KPI evaluation theory used here comes from RE-AIM~\cite{reaim_framework} which helps to judge whether the phase can scale through reach, adoption, implementation and maintenance. Additionally, Proctor et al.'s~\cite{proctor2011implementation} implementation outcomes are also used to justify specific KPI choices through the values of acceptability, feasibility, adoption and sustainability. RE-AIM informs the questions that should be asked for scalability, Proctor justifies the KPIs for answering those questions. These frameworks combined ensure that the GTM execution is not purely assessed through sales but by whether the product can actually be adopted and sustained in the target settings, providing a more data driven model and feedback loop to execute the GTM strategy of highest returns. KPIs are used first to validate execution at individual units, then to test whether the execution model can be repeated within the phase. Once the strategy has been successful across multiple sites there is approval for progression to the next phase.

\begin{table}[htbp]
    \centering
    \footnotesize
    \caption{GTM execution figures.}
    \label{tab:ent-gtm-execution-figures}
    \begin{tabular}{p{3.2cm}p{2.6cm}p{4.0cm}p{4.5cm}}
        \toprule
        \textbf{GTM stage} & \textbf{Timeframe} & \textbf{Target} & \textbf{Expected results used in SOM} \\
        \midrule
        Phase 1: Research validation & 12--18 months & 20--30 researchers / clinics contacted & 3--5 sites; 200--300 completed tests \\
        Phase 2A: School pilot & 1 school sport season & 60--80 sport-heavy schools contacted & 8--12 pilot schools \\
        Phase 2B: Youth club pilot & 1 rugby/football season & 60--70 clubs contacted & 6--8 pilot clubs \\
        Phase 2 paid institutional rollout & Following 12 months & 500--700 schools/clubs contacted & 70--100 paid institutions; base case 90 \\
        Phase 3 parent soft launch & First 3 months of Phase 3 & Parents exposed through 108 deployed schools/clubs from Phase 2 & 400--1,000 home users from approx. 20,000 exposed families \\
        Phase 3 parallel scaled rollout & First 12 months of Phase 3 & 800--1,200 further institutions contacted; 150 base-case paid conversions, 300 upside conversions & Contributes additional institutional revenue and year-weighted parent exposure for SOM \\
        \bottomrule
    \end{tabular}
\end{table}

\begin{table}[htbp]
    \centering
    \footnotesize
    \caption{Reference basis for the GTM execution figures.}
    \label{tab:ent-gtm-execution-rationale}
    \begin{tabular}{p{3.7cm}p{11.3cm}}
        \toprule
        \textbf{GTM stage} & \textbf{Reference basis / rationale} \\
        \midrule
        Phase 1: Research validation & Treated as a feasibility/validation deployment rather than a sales launch. NHS digital-health pilot guidance~\cite{ohid2020digitalhealth} supports pilots with defined scope, data-collection plans, success/failure criteria, fixed timelines and defined responsibilities. \\
        Phase 2A: School pilot & Parker et al.~\cite{parker2022schoolfeasibility} found UK school-based health feasibility trials recruited a median of 7.5 schools, with IQR 4.5--9 and range 2--37. This is used as a pilot-size comparator rather than UK commercial geography; 8--12 schools is a reasonable feasibility-scale pilot for the EU/EEA beachhead. \\
        Phase 2B: Youth club pilot & EU Six Nations rugby alone includes approximately 334,000 youth players across 2,577 clubs in France, Ireland and Italy~\cite{worldrugby2020participationmap}, so 6--8 clubs is a small practical feasibility sample from a much larger EU/EEA contact-sport pool. \\
        Phase 2 paid institutional rollout & Not literature-derived; this is a base-case commercial assumption conditional on pilot KPIs being met. Customer-development logic~\cite{blank2013startup} supports validating buyer, channel and repeatable sales process before scaling acquisition. \\
        Phase 3 parent soft launch & Modelled from institutional exposure rather than TAM. Rogers/Moore logic~\cite{rogers2003diffusion,moore1991chasm} supports using institutional proof, observability and trust before entering mainstream parent adoption. \\
        Phase 3 scaled rollout & Upside case depends on Phase 2 case studies, institutional testimonials and repeatable channel validation. Larger future volume is supported by continued European grassroots football/rugby participation and governing-body distribution routes~\cite{uefa2025schools,worldrugby2020participationmap}. \\
        \bottomrule
    \end{tabular}
\end{table}

\subsection{Phase 1}

The objective of phase 1 is to create a reliable product through iterations with researchers, creating credible validation evidence before later pilots. This phase is not focused on sales, and is centered on creating the best product and gathering validation from the researcher community to support CE/MDR-facing validation and EU/EEA institutional adoption in phase 2.

Therefore, the targets will include: 2--3 university sport science / neuroscience / engineering departments, 1 concussion clinic or sports medicine clinic and Oxford-linked researchers as warm trust channels are more effective.

The channels and tactics used in this phase will include: Oxford supervisor / academic introductions as researchers are more likely to respond through trusted academic channels; research-use-only headset loans to optimise partner acceptance rate and usability logging during every trial to capture product failure points and optimise the product's success rate and setup time.

The proposed timeline for this phase is 12--18 months and the phase-gate KPIs include 3--5 sites recruited, 200--300 completed tests, an acceptable test-failure rate, regulatory approval and a publishable paper for submitted validation output.

If these KPIs are not sufficiently achieved, iteration methods for improving them include: using warmer introductions and narrower research targets to help stand out in the noise~\cite{blank2013startup}; improving engineering workflows with potential engineering recruitment to produce better test-failure; spending more time conducting research trials to ensure a publishable output as evidence strength can still be insufficient after this much time.

The choices for these number of sites and tests is also supported in pilot-study literature for small feasibility samples~\cite{julious2005pilot}. Julious gives a rule of thumb for these numbers. Additional evidence of this strategy is given in the NHS pilot guidance~\cite{ohid2020digitalhealth} showing that pilots should have a defined scope, data-collection plan, success/failure criteria, fixed time period and defined responsibilities between the developer and research site.

These specific KPIs were selected to test whether the research-validation model is scalable beyond a single site. RE-AIM~\cite{reaim_framework} supports site recruitment to measure Reach and Adoption as well as test-failure rate measuring Implementation. Proctor~\cite{proctor2011implementation} justifies these metrics further: feasibility is measured through test-failure rate and acceptability through researcher feedback.

\subsection{Phase 2}

Schools and clubs will both be offered an institutional package of three headsets and three higher-tier app subscriptions, priced at \pounds{}300 per headset and \pounds{}600 per annual subscription as determined in \Cref{chapter:ent-financial,chapter:ent-business}. This gives \pounds{}2,700 first-year revenue per institution and \pounds{}1,800 annual renewal revenue. The objective for both is to ensure strong fit within their workflows, generate case studies as well as verifying channels for expansion within this segment whilst gaining trust for parent adoption and the mass market in Phase 3.

\subsubsection{2A -- Schools}

Referring back to the power-interest grids, targeted individuals include: headteachers who play a high-power, high-interest role and purchase the product; PE teachers who can assist in spreading the message due to their high interest and will convey the message of the product offering fast, simple, low-friction workflows after a head impact.

During pilot recruitment, 60--80 sport-heavy schools are contacted to secure 8--12 school pilots. If KPIs are met then the following paid rollout to schools form part of the wider 500--700 institutional contact list contributing to \textasciitilde{}50 in the base case. Tactics to optimise the execution success will include: recruiting in the pre-season/school planning window for the selected EU/EEA beachhead, provisionally May--August, before the autumn sport season to optimise the pilot recruitment rate; one-season free/discounted pilot as this will lower the barrier to adoption and increase the pilot conversion with repeated usage; targeting sport-heavy schools in selected EU/EEA beachhead regions first; using specific messaging with headteachers that ``This gives the school a defensible, documented workflow after a head impact'', as shown that this is a feature that matters from user interviews in \Cref{chapter:ent-adoption-roadmap}.

\subsubsection{2B -- Youth Clubs}

During pilot recruitment, 60--70 clubs are contacted to secure 6--8 club pilots. If the KPIs are met then in the following paid rollout, clubs contribute approximately 40 paid deployments within the 90-institution base case. Channels and tactics here will include: starting with EU/EEA rugby and football clubs and safeguarding/welfare-equivalent networks in the selected beachhead region, collecting anonymised match-day usage and targeting welfare or safeguarding-equivalent officers with evidence from the power interest grids showing that these are the key stakeholders to target and incentivise.

\subsubsection{Overall}

The timelines for both will include: recruitment in the selected EU/EEA beachhead's pre-season and school-planning window, provisionally May--August; September--March deployment, user interviews and failed-test analysis; June--July pilots converted into annual packages and recruiting the next cohort.

The pilot sizes were benchmarked against feasibility-study practice~\cite{parker2022schoolfeasibility}: UK school-based health feasibility trials recruit a median of 7.5 schools, which is used only as a pilot-size comparator and makes an 8--12 school pilot reasonable. For clubs, the EU Six Nations subset in \Cref{tab:ent-six-nations-rugby} provides a non-UK contact-sport base for 6--8 rugby pilots, while UEFA Football in Schools provides the parallel football route~\cite{uefa2025schools,worldrugby2020participationmap}. The 15--20 institution total and one-season timeline allow testing across both schools and clubs with defined scope, data collection and success criteria; the later 50--100 paid rollout is base-case commercial assumption conditional on pilot KPIs being met, not a literature-driven forecast.

KPIs here include clubs / schools recruited, pilot-to-paid conversion, tests per institution, successful test-completion rate, repeat usage across the season, PE teacher/coach confidence change, and renewal intent. These diagnose the effectiveness of the channel strength, workflow-feasibility, operator acceptability, social-pressure relief and real usage and whether the strategy can be repeated at scale. If weak, iteration techniques can include using warmer rugby networks, simplifying the access and storage of the headset and adjusting the message towards stakeholders; pushing more heavily into social proof tactics can improve adoption rates and will be a critical part of this, drawing on the behavioural economics evidence from \Cref{chapter:ent-market} and the worried parents in user interviews. These allow the testing of whether the model can be scaled in this phase and is also backed by RE-AIM and Proctor~\cite{reaim_framework,proctor2011implementation} as the number of institutions measures Reach as well as repeat usage / intent of renewal measuring sustainability.

Assuming these KPIs are hit, the first paid institutional rollout targets 70--100 paid deployments, with 90 total paid institutions used as a base case. Including the 15--20 pilots, this gives approximately 108 total deployed institutions before Phase 3. As a result, this will result in \pounds{}2,700 $\times$ 90 = \pounds{}243,000 pre-Phase 3 revenue. Phase 2 will also have a low paid-advertising spend: acquisition is driven through warm outreach, school networks, governing-body networks and pilot case studies due to this phase's main objective being generating enough trust to target parents and being aggressive expansion of Phase 2. These costs will not be deducted from the SOM as we define the SOM as gross obtainable revenue.

\subsection{Phase 3 -- Parents}

The objective here is to convert parents already exposed through schools and clubs into home users and expand to the mass market of schools, clubs and parents. Parents will be offered the product of the headset at \pounds{}300 with no subscription as determined in \Cref{chapter:ent-financial,chapter:ent-business}. Schools and youth clubs will still be targeted using previous strategies at a larger scale; for the SOM base case, this parallel rollout assumes 150 additional paid schools/clubs are added during the first Phase 3 year from a wider 800--1200 contact list; parents will be the new category of expansion here as we have gained sufficient evidence from the KPI metrics to progress to this phase.

Channels and tactics used here will include: emails to parents from schools and clubs referring back to the trust arguments of \Cref{chapter:ent-adoption-roadmap}; this will transfer trust from the schools and clubs to the product, assuring parents and improving the conversion rate KPI; home monitoring offers after an incident, offering the product to a parent whose child has already been affected increases product distribution and the repeat-test rates; creating clinician-ready/referral-ready PDF reports which can produce downstream value as optimising report downloads for this high-interest stakeholder incentivises them to advertise through word of mouth; screening aid, not diagnosis language persistently used to reduce ethical/regulatory misuse risks to reduce misuse/false reassurance reports. Parent messaging must be localised to the selected EU/EEA country, including school communication norms, GDPR and child-data consent requirements, and translated parent emails or app copy where the beachhead is non-English-speaking.

KPIs used here include: parent conversion from exposed institutions which analyses trust transfer; app onboarding completions, diagnosing product usability; repeat-test rate understanding longitudinal value; misuse/false reassurance reports to analyse ethical safety. If weak, iteration methods will include simplifying the app setup, adjusting email timings, increasing testimonial proof publicity and creating a clearer recovery-monitoring journey.

The timeline here will consist of the first 3 months having a soft launch to parents in pilot/paid institutions. If KPI metrics are hit, the next 9 months will consist of expansions to all families in Phase 2 institutions utilising the channels of schools/clubs emails to transfer the validity, pushing the adoption friction of trust stressed in \Cref{chapter:ent-adoption-roadmap}. KPI analysis is reinforced here to ensure optimal strategy before aggressive expansion. This phase will simultaneously have expansion of schools and clubs using strategies outlined in Phase 2 whilst starting the target of parents at home.


\section{Serviceable Obtainable Market}
\label{sec:ent-som}

SOM is defined as the expected obtainable revenue over the first 12 months after Phase 3 begins, assuming Phase 1 and Phase 2 KPIs have been met.

\begin{table}[htbp]
    \centering
    \footnotesize
    \caption{SOM revenue treatment by GTM period.}
    \label{tab:ent-som-periods}
    \begin{tabular}{p{3.4cm}p{6.8cm}p{4.0cm}}
        \toprule
        \textbf{Period} & \textbf{GTM activity} & \textbf{Revenue treatment} \\
        \midrule
        Phase 1: 12--18 months & Research validation with 3--5 sites and 200--300 tests & No revenue \\
        Phase 2 pilot: one season & 15--20 free/discounted EU/EEA school and club pilots & No revenue assumed \\
        Phase 2 paid rollout: 12 months & 90 paid EU/EEA institutions; 108 total deployed institutions including pilots & \pounds{}243,000 from 90 $\times$ \pounds{}2,700 \\
        Phase 3: first 12 months & Parent channel added while EU/EEA schools/clubs continue scaling & Main SOM window \\
        \bottomrule
    \end{tabular}
\end{table}

\begin{table}[htbp]
    \centering
    \footnotesize
    \caption{Institutional Phase 3 revenue.}
    \label{tab:ent-som-institutional}
    \begin{tabular}{p{7.4cm}p{3.2cm}p{2.8cm}}
        \toprule
        \textbf{Source} & \textbf{Calculation} & \textbf{Revenue} \\
        \midrule
        Existing paid EU/EEA schools/clubs renewing three app subscriptions & 90 $\times$ \pounds{}1,800 & \pounds{}162,000 \\
        New EU/EEA schools/clubs added during Phase 3 & 150 $\times$ \pounds{}2,700 & \pounds{}405,000 \\
        \midrule
        Total institutional Phase 3 revenue & & \pounds{}567,000 \\
        \bottomrule
    \end{tabular}
\end{table}

\begin{table}[htbp]
    \centering
    \footnotesize
    \caption{Effective parent exposure in Phase 3.}
    \label{tab:ent-som-family-exposure}
    \begin{tabular}{p{6.6cm}p{5.2cm}p{2.8cm}}
        \toprule
        \textbf{Source} & \textbf{Calculation} & \textbf{Effective exposed families} \\
        \midrule
        Families exposed from existing 108 deployed institutions & Approx. 20,000 & 20,000 \\
        Families exposed from 150 new Phase 3 institutions & 150 $\times$ \textasciitilde{}180 families $\times$ 50\% year weighting & 13,500 \\
        \midrule
        Total effective exposed families & & 33,500 \\
        Parent conversion and revenue & 33,500 $\times$ 3\% $\approx$ 1,000 users; 1,000 $\times$ \pounds{}300 & \pounds{}300,000 revenue \\
        \bottomrule
    \end{tabular}
\end{table}

Each school is modelled as exposing 200 families as a retained EU/EEA beachhead assumption from the current SOM model. This is broadly in line with the simple average implied by UEFA Football in Schools (7.2 million children across 41,500 schools, or approximately 173 children per school before country mix, siblings and multi-child households are adjusted)~\cite{uefa2025schools}, and should be validated against the selected country during Phase 2. Each club is 150 families as a conservative grassroots-club modelling assumption~\cite{ukparliament2023clubs}. Therefore 60 $\times$ 200 + 48 $\times$ 150 = 19,200 rounds up to 20,000. Numbers are used again for Phase 3 exposure too. The 3\% conversion is a base case assumption due to the absence of a directly comparable warm-channel benchmark for parent-facing health screening and is made more plausible by the fact that this is not cold advertising; parents are exposed through a trusted channel after institutional validation. To reflect this uncertainty, the SOM includes a conservative 2\% case and an upside 5\% case.


\begin{table}[htbp]
    \centering
    \footnotesize
    \caption{SOM scenario sensitivity.}
    \label{tab:ent-som-scenarios}
    \begin{tabular}{lrrrrrr}
        \toprule
        \textbf{Scenario} & \textbf{Existing} & \textbf{New} & \textbf{Institutional} & \textbf{Exposed} & \textbf{Parent} & \textbf{Phase 3} \\
        & \textbf{paid} & \textbf{Phase 3} & \textbf{revenue} & \textbf{families} & \textbf{revenue} & \textbf{SOM} \\
        \midrule
        Conservative & 50 & 75  & \pounds{}292,500  & 16,750 & \pounds{}100,500 & \pounds{}393,000 \\
        Base case    & 90 & 150 & \pounds{}567,000 & 33,500 & \pounds{}300,000 & \pounds{}867,000 \\
        Upside       & 150 & 300 & \pounds{}1,080,000 & 62,000 & \pounds{}930,000 & \pounds{}2,010,000 \\
        \bottomrule
    \end{tabular}
\end{table}

Phase 1 generates no revenue because it is a UK research-validation phase, while Phase 2 pilots are assumed to be free or discounted. Phase 2 produces 90 paid EU/EEA institutional deployments and 108 total deployed sites, generating \pounds{}243,000 from the initial paid institutional deployment. In Phase 3, the institutional channel continues rather than stopping: the base case assumes 150 additional paid EU/EEA schools/clubs are added, while the 90 existing paid institutions renew three \pounds{}600 annual subscriptions each. This produces \pounds{}567,000 of institutional revenue in the first Phase 3 year. Parent SOM is calculated from families exposed through deployed institutions. The existing 108 institutions expose approximately 20,000 families; new Phase 3 institutions are year-weighted at 50\% assuming uniform onboarding throughout the year, adding an effective 13,500 exposed families. At a 3\% parent conversion rate and \pounds{}300 per headset, this gives approximately 1,000 parent users and \pounds{}300,000 parent revenue. Therefore, the base-case first-year Phase 3 SOM is \pounds{}867,000. Conservative and upside scenarios are included to reflect uncertainty in institutional conversion and parent trust transfer.

\section{Chapter Conclusion and Limitations}
This chapter has developed an evidence-based execution plan for commercialising the headset-based mTBI screening system. Building on the adoption sequence justified in \Cref{chapter:ent-adoption-roadmap}, the GTM Execution Strategy section turned that sequence into an execution plan using critical KPIs for validation, and the Serviceable Obtainable Market section then took these GTM assumptions and produced a bottom-up SOM of base-case \pounds{}867,000 with conservative and upside scenarios to consider uncertainties. This base-case SOM represents \textasciitilde{}0.04\% of the \pounds{}2.27 billion conservative European school-football SAM in year 1.
Although this structure is laid out with flexibility for iteration, it is very possible that the data shows that the final strategy is much different to what was predicted, and that parents may adopt earlier than expected without having to do an entire year of Phase 2. Simultaneously, we may also find that conversion rates are poor and that the product fails to cross the chasm, with evidence showing that headset technologies face technical, user-interface, and social-acceptance barriers, primarily for head-mounted systems ~\cite{billinghurst2015ar} and crossing this chasm is a known high-risk point~\cite{moore1991chasm}.
